---
title: "Assignment 04 - HPC and SQL"
author: "Brandyn Ruiz"
date: "10/24/2020"
output: github_document
---

The learning objecives are to conduct data scraping and perfrom text mining.

```{r, message=FALSE, echo=TRUE, warning=FALSE}
library(parallel)
library(RSQLite)
library(DBI)
```

# HPC

## Problem 1: Make sure your code is nice

Rewrite the following R function to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google.

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE
  ans <- NULL
  for (i in 1:nrow(mat)){
    ans[i] <- sum(mat[i, ])
  }
  ans
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
  ans <- mat
  for(i in 1:nrow(mat)){
    for(j in 2:ncol(mat)){
      ans[i, j] <- ans[i, j] + ans[i, j - 1]
    }
  }
  ans
}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```

The last arguement, check = "equivalent", is included to make suree that the functions return the same result.

## Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using `parLapply()` to make it run faster. Make sure you set the seed using `clusterSetRNGStream()`:

```{r}
# YOUR CODE HERE
cl <- makePSOCKcluster(4)
x <- 4000
clusterSetRNGStream(cl, 1231) # Equivalent to `set.seed(1231)`

system.time({
  # YOUR CODE HERE
  ans <- NULL
  print(mean(ans))
  # YOUR CODE HERE
  parallel::stopCluster(cl)
})
```

# SQL
Setup a temporary datavase by running the following chunk

```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv(
  "https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")

film_category <- read.csv(
  "https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")

category <- read.csv(
  "https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

When you write a new chunk, remember to replace the `r` with `sql, connection = con`. Some of these questions will require you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp

## Question 1
How many movies is there avaliable in each `rating` category?

```{sql connection=con}
SELECT rating, COUNT(*) as count
FROM film
GROUP BY rating
ORDER BY count DESC
```

From our output above we see that there are more movies with the rating of PG-13 with a count of 223, following is NC-17 with 210 total, rated R movies have a value of 195, PG movies with 194, and rated G 180 movies.

## Qustion 2
What is the average replacement cost and rental rate for each `rating` category?

```{sql connection=con}
SELECT rating, avg(replacement_cost) as AVGreplacement_cost, avg(rental_rate) as AVGrental_rate
FROM film
GROUP BY rating
ORDER BY AVGreplacement_cost DESC, AVGrental_rate DESC
```

The greatest movie rating with the highest average replacement cost and renatal rate are PG-13 movies. The least average replacement cost is $18.96 with the lowest average rental rate of $3.05 for PG movies.

## Question 3
Incorporate table `category` into the answer to the previous question to find the name of the most popular category

```{sql connection=con}
SELECT film.rating, avg(film.replacement_cost) as AVGreplacement_cost,
  avg(film.rental_rate) as AVGrental_rate, category.name
FROM film
JOIN film_category ON film.film_id = film_category.film_id
JOIN category ON category.category_id = film_category.category_id
GROUP BY rating
ORDER BY AVGreplacement_cost DESC, AVGrental_rate DESC
```

With incorporating the category table its contents have `category_id` and `name`, but in order to connect the `category_id` to the *film* table our *film_category* table must also be involved. From *film_category* we can connect the two datasets with `film_id` and `category_id`. When arranging the averages of the replacement cost and average rental rate in descending order, movie rating with PG-13 being Comedy is the greatest, while PG Documentary movies are the least.